Lesson 4
========================================================

### Scatterplots
Notes: we use scatterplot to visualize the relationship between two continuous variables.

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep ='\t')
qplot(x= age, y= friend_count, data= pf)
```

***

#### What are some things that you notice right away?
Response: Younger users have a lot of friends.

***

### ggplot Syntax
Notes: ggplot let us specify more complicated plots.

```{r ggplot Syntax}
ggplot(aes(x= age, y= friend_count), data= pf)+
  geom_point() + xlim(13, 90)

summary(pf$age)
```

***

### Overplotting
Notes: overplotting makes it different to tell how many points are in each region, so we can set the transparency of the points using the alpha parameter and geom point.

```{r Overplotting}
ggplot(aes(x= age, y= friend_count), data= pf)+
  geom_jitter(alpha=1/20) +
  xlim(13, 90)
```

#### What do you notice in the plot?
Response: The friend count for young users aren't nearly as high as they looked before, the bulk of young users really have friend counts below 1000.

***

### Coord_trans()
Notes: with this plot, it's much easier to see the distribution of friend count.

```{r Coord_trans()}
ggplot(aes(x= age, y= friend_count), data= pf)+
  geom_point(alpha=1/20, position= position_jitter(h=0)) +
  xlim(13, 90)+
  coord_trans(y= 'sqrt')
```


***

### Alpha and Jitter
Notes: we use alpha to reduce overplotting

```{r Alpha and Jitter}
ggplot(aes(x= age, y= friendships_initiated), data= pf)+
  geom_point(alpha= 1/20, position= position_jitter(h=0))+
  coord_trans(y='sqrt')
```

***

### Conditional Means
Notes: Created another data frame that contains the mean and the median of friend count for each age.

```{r Conditional Means}
library("dplyr")

pf.fc_by_age <- 
  pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n<- n()) %>%
  arrange(age)

head(pf.fc_by_age, 20)
```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x= age, y= friend_count_mean), data= pf.fc_by_age) +
  geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes: displaying multiple summaries at the same time on the plot.

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x= age, y= friend_count), data= pf)+
  xlim(13, 90) +
  geom_point(alpha= 0.05,
             position= position_jitter(h = 0),
             color= 'orange')+
  coord_trans(y = 'sqrt') +
  geom_line( stat= 'summary', fun.y= mean) +
  geom_line(stat= 'summary', fun.y= quantile,
            fun.args= list(probs = .1),
            linetype= 2, color= 'blue') +
  geom_line(stat= 'summary', fun.y= quantile,
            fun.args= list(probs = .5),
            color= 'blue') +
  geom_line(stat= 'summary', fun.y= quantile,
            fun.args= list(probs = .9),
            linetype= 2, color= 'blue')
```

#### What are some of your observations of the plot?
Response: more than 1000 friend is rare, 90% have less than 1000 friends.

***

### Correlation
Notes: We used the Pearson product moment correlation to measure the linear relationship between age and friend count.

```{r Correlation}
cor.test(pf$age, pf$friend_count, method= 'pearson')
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response: -0.027

***

### Correlation on Subsets
Notes: We don't want to include the older ages in our correlation number, since older ages are likely to be incorrect.

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes: Correlation Methods: Pearson's r, Spearman's rho, and Kendall's tau.

***

## Create Scatterplots
Notes: We will look at the number of likes users recieved from friends on the desktop version on the site and compare it to the total number of likes users recieved.

```{r}
ggplot(aes(x= www_likes_received, y= likes_received), data = pf) +
  geom_point()
```

***

### Strong Correlations


```{r Strong Correlations}
ggplot(aes(x= www_likes_received, y= likes_received), data = pf) +
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95))+
  ylim(0, quantile(pf$likes_received, 0.95)) +
  geom_smooth(method = 'lm', color = 'red')
```


```{r Correlation Calcuation}
cor.test(pf$www_likes_received, pf$likes_received)
```

***

### More Caution with Correlation

```{r More Caution With Correlation}
library(alr3)
data("Mitchell")
```

Create your plot!

```{r Temp vs Month}
ggplot(data = Mitchell, aes(x= Month, y= Temp)) +
  geom_point()
```

***

### Noisy Scatterplots


```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes: We should break months into 12 as we have 12 months in the year

```{r Making Sense of Data}
ggplot(data = Mitchell, aes(x= Month, y= Temp)) +
  geom_point() +
  scale_x_continuous(breaks = seq(0, 203, 12))
```

***

### A New Perspective

What do you notice?
Response: When stretching out of the graph, we notice that we get more of a cyclical pattern because there are seasons in Nebraska.

***

### Understanding Noise: Age to Age Months

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) +
  geom_line()

pf$age_with_months <- pf$age + (12- pf$dob_month)/12

```

***

### Age with Months Means

```{r Age with Months Means}
pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age_with_months)

head(pf.fc_by_age_months)
```


***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(x = age_with_months, y = friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months <71)) +
  geom_line()
```

***

### Smoothing Conditional Means

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x = age, y = friend_count_mean), data = subset(pf.fc_by_age, age <71)) +
  geom_line()+
  geom_smooth()

p2 <- ggplot(aes(x = age_with_months, y = friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months <71)) +
  geom_line()+
  geom_smooth()

p3 <- ggplot(aes(x = round(age / 5) * 5, y = friend_count), data = subset(pf, age <71)) +
  geom_line(stat = 'summary', fun.y = mean)

library(gridExtra)
grid.arrange(p2, p1, p3, ncol = 1)
```

***

### Which Plot to Choose?
Notes: In exploratory data analysis we often create multiple visualizations and summaries of the same data, gleaning different incites from each.

***


Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

