Lesson 5
========================================================

```{r}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep ='\t')
```

***

### Third Qualitative Variable

```{r Third Qualitative Variable}
library(dplyr)

pf.fc_by_age_gender <- 
  pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n<- n()) %>%
  ungroup() %>%
  arrange(age)

head(pf.fc_by_age_gender)
```

***

### Plotting Conditional Summaries

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = friend_count_median),
       data = pf.fc_by_age_gender) +
  geom_line(aes(color = gender))
```

***

```{r}
#install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'friend_count_median')

head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot

```{r Ratio Plot}
ggplot(aes(x = age, y = female/male),
       data = pf.fc_by_age_gender.wide) +
  geom_line() +
  geom_hline(yintercept = 1, alpha = 0.3, linetype = 2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)
```

***

### Cut a Variable
Notes:
the cut function is useful for making discrete variables from continuous or nomerical ones.
```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014))
```

***

### Plotting it All Together

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes: plot the mean instead of the median

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)
```

***

### Friending Rate
Notes:
we want to see how many friends does a user have for each day since starting using facebook.

```{r Friending Rate}
with(subset(pf, tenure >= 1), summary(friend_count/tenure))
```

***

### Friendships Initiated

What is the median friend rate? 0.2205

What is the maximum friend rate? 417

```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Bias-Variance Tradeoff Revisited
Notes: we can adjust the noise from the last graph by bending our x-axis differently.

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Introducing the Yogurt Data Set
Notes: the facebook data has one row per individual, on the other hand, Yogurt data has many rows per household, one for each purchase occasion.

***

### Histograms Revisited
Notes: First, convert id from integer into factor. Second, create a histogram of the Yogurt prices.

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
str(yo)

#change the id from int to a factor
yo$id <- factor(yo$id)
str(yo)

qplot(x = price, data = yo, fill = I('#F79420'))
```

***

### Number of Purchases
Notes: all.purchases variable gives the total counts of yogurt for each observation or household purchase.

```{r Number of Purchases}
summary(yo)
table(yo$price)

yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)

summary(yo$all.purchases)
```

***

### Prices over Time

```{r Prices over Time}
ggplot(aes(x= time, y= price),
       data= yo) +
  geom_jitter(alpha = 1/4, shape= 21, fill = I('#F79420'))
```

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x= time, y= price),
       data= subset(yo, id %in% sample.ids)) +
  facet_wrap( ~ id) +
  geom_line() +
  geom_point(aes(size = all.purchases), pch = 1)
```

***

### Scatterplot Matrix

```{r}
#install.packages('GGally')
library('GGally')
theme_set(theme_minimal(20))

set.seed(1836)
pf_subset <- pf[, c(2:15)]
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ])

```

***

### Heat Maps

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***


Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

